Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/Response-MajRev/response_letter.tex

% !TEX program=pdflatex
% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{times}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{float}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{soul}
\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
% Vector
\newcommand{\MyVec}[1]{\boldsymbol{#1}}
% Matrix
\newcommand{\MyMat}[1]{\boldsymbol{#1}}
% Vectors with hat
\newcommand{\hvec}[1]{\hat{\boldsymbol{h}}_{#1}}
% Model parameters
\newcommand{\Param}[1]{\textcolor{blue}{#1}}

@@

TODO

% Let's start point-by-point with Reviewer 1
\reviewersection

% R1Q0
\begin{point}
    The authors conduct an empirical analysis of performance bottlenecks in graph neural network training. The authors identify the edge-related calculation is the performance bottleneck. Experimental of several GNN variants, such as GCN, GGNNNN, GAT and GaAN on six real-world graph datasets verify the importance of the findings. The experimental analysis is sufficient. However, there are some tiny issues in this paper.
\end{point}

\begin{reply}
    Thank you for your kindly comments on our manuscript.
    %
    We have carefully revised the manuscript according to your suggestions. 
    %
    Please see the replies below.
    %
    The changes corresponding to each point are highlighted by red squares in the annotated version of our revised manuscript.

@@

     \begin{figure}[H]
        \centering
        \subfloat[Batch Size = 1024]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_1024.pdf}}
        \subfloat[Batch Size = 2048]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_2048.pdf}}
        \caption{Inference time per batch breakdown under different batch sizes.}
        \label{fig:time_breakdown_of_inference_sampler}
     \end{figure}
\end{reply}

% Begin a new reviewer section
\reviewersection

\begin{point}
    In the paper, authors accomplished a unique study and analysis on GNN models training complexity.  The articles first review and development history of GNNs and creatively model all architectures as input layers, intermediate layers of graph neurons and prediction layers. And they quantitatively summarize the time and space complexity of 4 representative GNNs, including graph convolution, gated recurrent graph net, graph attention net and GraphSage. Most importantly, the article first break down complexity into operator level and offered analysis of good granularity, giving reader more guidance in future study. At last, the solid experiments included the study of effects of hyper-parameters and a comparison of two major sampling techniques: neighbor sampling and cluster sampling.
\end{point}

\begin{reply}
    Thank you for your positive comments on our manuscript.
    %
    We have carefully revised the manuscript according to your kindly comments and suggestions in the following points.
    %
    We highlight our modifications with red squares point by point in the annotated version of the revised manuscript.
\end{reply}

@@

\begin{point}
    In general, the paper was well written and organized with good structure and clear narratives. Just some minor language errors like line Page 8, Line 208, "In active graph neurons" =\textgreater "Inactive graph neurons".
\end{point}

% R2Q1
\begin{reply}
    We feel really sorry for our carelessness.
    %
    We have proofread our revised manuscript to eliminate such language errors as much as we can.
\end{reply}

@@

    And the experimental evaluation were done over 6 large graph-structured datasets.
\end{point}

\begin{reply}
    Thank you very much for your appreciation.	
\end{reply}

% R2Q3
\begin{point}
    While, one major drawback is that I did not clearly see the analysis complexity v.s. accuracy. For example, in Figure 19 and 20, I did not see network accuracy from those 4 GNNs. There is always tradeoff between model complexity and model performance, and in some scenarios where high complexity is allowed, a sophisticated model of more powerful representation capability is still needed.
\end{point}

\begin{reply}
    Thanks very much for your valuable suggestions.
    %
    The model complexity directly affects both the accuracy and the training time.
    %
    In order to analyze the relationship between model complexity and accuracy, we have conducted two kinds of extra experiments in the revised manuscript: (1) how the hyper-parameters of the GNNs (like the dimension of hidden vectors and the number of heads) affect the accuracy of GNNs (in Section 4.1); (2) how the batch size in the sampling methods affects the accuracy of GNNs (in Section 4.4).

    We have added more dataset description related to accuracy evaluation in Section 3.1 ``Experimental Setting''.
    %
    Newly added sentences are marked with underlines in the following quotation.

    \begin{myquote}{Section 3.1 ``Experimental Setting''}
    \emph{Datasets}.
    %
    We used six real-world graph datasets as listed in \tablename~\ref{tab:dataset_overview} (in the reply) that were popular in the GNN evaluation [28,31,34,36].
    %
    For directed graphs, PyG converts them into undirected ones during data loading.
    %
    Thus, the average degree of a directed graph $\bar{d}=\frac{2|\mathcal{E}|}{|\mathcal{V}|}$.
    %
    For an undirected graph, the average degree is defined as $\bar{d}=\frac{|\mathcal{E}|}{|\mathcal{V}|}$.
    %
    \ul{For the \texttt{cam} dataset, its vertices were not associated with feature vectors}.
    %
    \ul{Thus, we generated random dense feature vectors for it and excluded it from accuracy evaluation}.
    %
    \ul{For \texttt{amp}, \texttt{amc}, \texttt{cph}, and \texttt{cam}, we used 70\%/15\%/15\% of the vertices as the training/evaluation/test set, respectively}.
    %
    \ul{For \texttt{pub}, we used 500/1000 vertices as the evaluation/test set and the remained as the training set according to [28]}.
    %
    \ul{For \texttt{fli}, we used 50\%/25\%/25\% of the samples as the training/evaluation/test set according to [31]}.
    %
    We also used random graphs generated by the R-MAT graph generator [35] in some experiments, to explore the effects of graph topological characteristics (like average degrees) on performance bottlenecks.
    %
    Input feature vectors of random graphs were random dense vectors with the dimension of 32.
    %
    Vertices of random graphs were classified into 10 classes randomly.

    \begin{table}[H]
        \centering
        \begin{tabular}{cccccccc}
            \toprule
            Dataset                                                 & $|\mathcal{V}|$ & $|\mathcal{E}|$ & $\bar{d}$ & $dim(\boldsymbol{v})$ & \#Class & Directed \\
            \midrule
            pubmed (\texttt{pub}) [28]  & 19,717          & 44,324          & 4.5       & 500                   & 3       & Yes      \\
            amazon-photo (\texttt{amp}) [34]     & 7,650           & 119,081         & 31.1      & 745                   & 8       & Yes      \\
            amazon-computers (\texttt{amc}) [34] & 13,752          & 245,861         & 35.8      & 767                   & 10      & Yes      \\
            coauthor-physics (\texttt{cph}) [34] & 34,493          & 247,962         & 14.4      & 8415                  & 5       & Yes      \\
            flickr (\texttt{fli}) [31]                 & 89,250          & 899,756         & 10.1      & 500                   & 7       & No       \\
            com-amazon (\texttt{cam}) [36]               & 334,863         & 925,872         & 2.8       & 32                    & 10      & No       \\
            \bottomrule
        \end{tabular}
        \caption{Dataset overview. $\bar{d}$ represents the average vertex degree. $dim(\boldsymbol{v})$ is the dimension of the input feature vector.}
        \label{tab:dataset_overview}
    \end{table}

    \end{myquote}

    In this reply, we focus on the experiments related to hyper-parameters in Section 4.1.
    %
    We introduce the experimental results of the sampling techniques in Section 4.4 in the next reply.

    Hyper-parameters determine the model complexity of a GNN.
    %
    Generally speaking, higher values of hyper-parameters bring higher model complexity and increase representation capability.
    %
    In the revised manuscript, we add a new subsection Section 4.1.4 ``Effects on Accuracy'' at the end of Section 4.1 ``Effects of Hyper-parameters on Performance'' to analyze how hyper-parameters affect the accuracy.
    %
    We have two main findings.
    %
    First, the accuracy of GNNs is much more sensitive to the dimension of hidden vectors $dim(\MyVec{h}^1_x)$ (for GCN/GGNN/GaAN) and the dimension of each head $d_{head}$ (for GAT) than the other hyper-parameters.
    %
    Second, the relative accuracy of the four typical GNNs varies greatly with different datasets.
    %
    We quote from Section 4.1.4 of the revised manuscript to present our results.

    \begin{myquote}{Section 4.1.4 ``Effects on Accuracy''}
    %
    The values of hyper-parameters determined the model complexity of a GNN.
    %
    The relationship between model complexity and accuracy was complex.
    %
    %As far as we know, there is little theoretical analysis yet.
    %
    Generally speaking, higher model complexity brought more powerful representation capability and might bring higher accuracy, but it also increased the risk of overfitting.

    To evaluate the effects of hyper-parameters on accuracy, we measured the accuracy of the typical GNNs with varying hyper-parameters.
    %
    \figurename~\ref{fig:exp_hyperparameter_accuracy} (in the reply) shows the experimental results.
    %
    For GCN, its accuracy was sensitive to the dimension of hidden vectors $dim(\MyVec{h}^1_x)$.
    %
    As $dim(\MyVec{h}^1_x)$ increased, the accuracy first increased quickly and then stabilized when $dim(\MyVec{h}_x^1) \geq 8$.
    %
    For GGNN, its accuracy curves showed similar trends as GCN, but GGNN was more sensitive to $dim(\MyVec{h}_x^1)$ than GCN.
    %
    Its accuracy even decreased when $dim(\MyVec{h}_x^1) \geq 1024$.
    %
    Since GGNN had high model complexity (with 13 weight matrices/vectors to train), GGNN might occur overfitting in those cases.
    %
    For GAT, its accuracy was more sensitive to the dimension of each head $d_{head}$ than the number of heads.
    %
    For GaAN, only $dim(\MyVec{h}_x^1)$ showed obvious impacts on accuracy.
    %
    The experimental results indicated that the accuracy of the GNNs was often low when $dim(\MyVec{h}_x^1)$ (for GCN/GGNN/GaAN) or $d_{head}$ (for GAT) was too low.
    %
    %The low model complexity limited the representation capability of the GNNs.
    %
    As $dim(\MyVec{h}_x^1)$ or $d_{head}$ increased to a certain threshold, the GNNs gained sufficient learning ability to achieve stable accuracy.

@@

    \begin{figure}[H]
        \centering
        \subfloat[GCN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gcn.pdf}}
        \subfloat[GGNN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_ggnn.pdf}}\\
        \subfloat[GAT]{\includegraphics[width=0.8\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gat.pdf}}\\
        \subfloat[GaAN]{\includegraphics[width=1.0\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gaan.pdf}}
        \caption{Effects of hyper-parameters on accuracy of the typical GNNs.}
        \label{fig:exp_hyperparameter_accuracy}
    \end{figure}

    %\begin{figure}[H]
        %\centering
        %\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat_small_info.pdf}
        %\caption{Effects of the number of heads on GAT ($d_{head}$=2,4)}
        %\label{fig:exp_hyperparameter_accuracy_gat_small_info}
    %\end{figure}

    \figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} (in the reply) further compares the best accuracy that each GNN could achieve on different datasets.
    %
    The best accuracy of the typical GNNs was very close.
    %
    %
    It was also close to the accuracy reported in their original references [34, 31].
    %
    The results indicated that there was no clear winner.
    %
    The relative accuracy between GNNs varied greatly with different datasets.
    %
    GaAN achieved the highest accuracy in three out of five datasets.
    %
    GCN achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
    %
    Simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

    \end{myquote}
\end{reply}

\begin{point}
    Sampling method is definitely going to reduce model complexity, since all models complexity depend on graph node number N, while performance is going to be compromised as well. I would like to see authors resolve the concern of significant accuracy drop after applying aggressive sampling of subgraphs.
\end{point}

\begin{reply}

    Thanks for the insightful comment and valuable suggestion.
    %
    The sampling methods indeed affect the test accuracy of GNN models [5, 28, 26, 31].
    %
    Without sampling, the GNN models are trained with a \emph{batch} gradient descent manner.
    %
    With sampling, the GNN models are trained with \emph{mini-bach} stochastic gradient descent.
    %
    The structure of the GNN remains unchanged between batches.
    %
    Each GNN layer always consists of $|\mathcal{V}|$ graph neurons, where $|\mathcal{V}|$ is the number of vertices in the graph.
    %
    But the numbers of \emph{activated} graph neurons are different for different batches.
    %
    In each batch, only the neurons corresponding to the vertices appearing in the sampled subgraph will be activated.
    %
    The model parameters are updated according to the gradients calculated on the sampled subgraph in each batch.
    %
    Thus, the test accuracy of the models trained with sampling techniques may be different from the full-graph training.

    To discuss how the sampling techniques affect model accuracy, we have added a new subsection Section 4.2.2 ``Effects on Accuracy'' in Section 4.4 ``Effects of Sampling Techniques'' in the revised manuscript.
    %
    We quote from Section 4.4.2 of the revised manuscript to present our results.

@@

    \begin{myquote}{Section 4.4.2 ``Effects on Accuracy''}

    Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models got by sampling techniques might be different from the full-batch training.
    %
    To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
    %
    \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} (in the reply) and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} (in the reply) compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
    %
    For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.

### 2020-12-25 NEUROC/sec2-review-of-graph-neural-network.tex
新增精简版的response letter.
更新了inference介绍和Inference阶段的时间性能瓶颈分析

The model parameters of a GNN are stored in the weight vectors/matrices of the prediction layer and the messaging/aggregation/updating functions of each GNN layer.
%
When we want to perform inference on a new graph based on a trained GNN, the structure of the GNN needs to be adjusted, but the hyper-parameters and the model parameters of the GNN remain unchanged.
%
The number of graph neurons in each GNN layer has to be adjusted to $|\mathcal{V}|$ of the new graph.
%
The connections of graph neurons between GNN layers also needs to be adjusted according to the edge set of the new graph.

During inference, input feature vectors of the new graph are propagated forward from the input layer to the prediction layer to give out predicted labels.
%
The workflow of inference is the same as the workflow of the \emph{forward} phase during training.
%
The only difference from the forward phase is that there is no need to cache intermediate calculation results during inference, because inference does not need to prepare for gradient calculation.

Inference methods can be divided into two types based on the range of input graphs involved in inference: full-batch inference and sample-based inference.
%
The full-batch inference performs inference on the whole input graph at once.
%
It can make full use of the computing power of GPUs, bringing high computation efficiency.
%
The full-batch inference is suitable for small graphs that the GPU memory can hold all feature vectors, models, and intermediate results.

Different from the full-batch inference, the sample-based inference only predicts labels for part of vertices/edges each time, limiting the GPU memory usage during inference.
%
It first samples a subgraph related to inference from the input graph and then performs the full-batch inference on the sampled subgraph.
%
The sampling method depends on the prediction task.
%
Taking the node classification scenario as an example, if there are $L$ GNN layers in the GNN, the hidden vector outputted by the last GNN layer of each vertex $\MyVec{h}^L_x$ is only related to the $L$-hop neighborhood of $v_x$.
%
Thus, if we want to predict labels for a group of vertices, we only need to sample a subgraph that contains $L$-hop neighborhoods of all given vertices.
%
The sample-based inference is suitable for processing big graphs.
%
It stores the whole input graph and its feature vectors in the main memory.
%
It processes the whole graph in batches.
%
The subgraph of each batch is sampled on the CPU side and sent to the GPU side to perform inference.
%
By using small batch sizes, the peak memory usage during inference can be limited under the GPU memory capacity.
%
The sample-based inference is also suitable for the online inference scenario that the topology of the graph evolves over time.
%
When a graph update arrives, the subgraph affected by the update is sampled and inferences are only performed on the sampled subgraph.it. that are affected by , avoiding conducting repeated inferences for the whole graph.

>>>>

The model parameters of a GNN consist of the weight vectors/matrices in the prediction layer and the messaging/aggregation/updating functions of each GNN layer.
%
For transductive learning, the input graphs of training and inference are same.
%
The structure of the GNN does not need to adjust. 
%
For inductive learning, the input graphs used in training and inference are different.
%
When we want to perform inference on a new graph with a pre-trained GNN, the structure of the GNN needs to be adjusted, but the hyper-parameters and the model parameters of the GNN remain unchanged.
%
The number of graph neurons in each GNN layer has to be adjusted to $|\mathcal{V}|$ of the new graph.
%
The connections of graph neurons between GNN layers are also adjusted according to the edge set of the new graph.
%
During inference, input feature vectors of the input graph are propagated forward from the input layer to the prediction layer to give out predicted labels.

When the memory capacity of the GPU is large enough to hold all of the input graph and intermediate results during inference, the inference can produce predicted labels for all vertices and edges at once, making full use of the computing power of the GPU.
%
When the memory capacity is not large enough, the sampling technique must be adopted.

The sample-based inference splits the test set into several batches.
%
For each batch, the sample-based inference samples a subgraph related to the batch from the input graph.
%
It performs inference on the sampled subgraph to produce predicted labels for the batch.
%
Since the sampled subgraph is usually much smaller than the original input graph, the memory usage during inference is limited.
%
The sampling method depends on the prediction task.
%
Taking the node classification task as an example, if there are $L$ GNN layers in the GNN, the hidden vector outputted by the last GNN layer of each vertex $\MyVec{h}^L_x$ is only related to the $L$-hop neighborhood of $v_x$.
%
Thus, if we want to predict labels for a group of vertices, we only need to sample a subgraph that contains $L$-hop neighborhoods of all given vertices.

The implementation of sample-based inference in PyG stores the whole input graph and its feature vectors in the main memory.
%
In the node classification task, it splits the vertices in the test set into batches accoding to a given batch size.
%
The subgraph corresponding to each batch is sampled on the CPU side and sent to the GPU side to perform inference.
%

### 2020-12-25 NEUROC/sec4-experiments.tex
更新了inference介绍和Inference阶段的时间性能瓶颈分析

\paragraph{Effects on Training Time}

@@

\paragraph{Effects on Inference Time}

@@

\paragraph{Effects on Peak Memory Usage}

@@

\paragraph{Effects on Accuracy}

@@

\subsection{Training Time Breakdown}
\label{sec:training_time_breakdown}

@@

% added by wangyunpan, 2020-12-22 ----start----
\paragraph{Inference Time Breakdown Analysis}
We conducted the same experiments in the inference stage as above.
%
The results of layer level experiment, step Level in edge calculation experiment and operator Level experiment were in general consistent with those in the Training stage.
%
The performance bottleneck analysis in the Training stage can also be applied to the inference stage.
% added by wangyunpan, 2020-12-22 ----end----

>>>>

\subsubsection{Effects on Training Time}

@@

\subsubsection{Effects on Inference Time}

@@

\subsubsection{Effects on Peak Memory Usage}

@@

@@

\subsection{Training/Inference Time Breakdown}
\label{sec:training_time_breakdown}

@@

\subsubsection{Performance Bottlenecks in Inference}
%
We also did all the above experiments for Inference.
%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Here, we used GCN as an example. 
%
We made similar observations for all GNNs. 
%
\figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
Compared to the training time breakdown of GCN \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.

% GCN Inference结果对比图
\begin{figure}[H]
    \centering
    \subfloat[\emph{layer} level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[the effect of average degree]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[\emph{edge} level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
    \caption{Time breakdowns on the layer level and edge level, and the effect of average degree on the edge/vertex calculation of GCN Inference.}
    \label{fig:inference_analysis_on_layer_edge_level_of_gcn}
\end{figure}

% 绝对时间对比图
\begin{figure}[H]
    \centering
    \subfloat[\texttt{amc}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-computers.pdf}}
    \subfloat[\texttt{amp}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-photo.pdf}}
    \subfloat[\texttt{fli}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_flickr.pdf}}
    \caption{Wall-clock training/inference time on different datasets.}
    \label{fig:compare_wall_clock_time_of_training_and_inference}
\end{figure}

The main differences between training and inference were reflected in two aspects: the wall-clock time and the top time-consuming basic operators.
%
The inference time was much less than the training time.
%
Figure~\ref{fig:compare_wall_clock_time_of_training_and_inference} in the response (Figure X in the revised manuscript) compares the wall-clock time of training and inference on the \texttt{amp}, \texttt{amc}, and \texttt{fli} datasets.
%
The results on the other datasets were similar.
%
Since the inference only conducted the forward propagation from the input layer to the prediction layer, the time of inference was very close to the time of the forward phase in training.
%
The inference time was 34\% (GCN), 32\% (GGNN), 25\% (GAT), and 32\% (GaAN) of the training time, averaged over all datasets.

The top time-consuming basic operators of training and inference also showed a certain degree of difference.
%
Some of the top time-consuming operators during training were replaced by new operators in inference.
%
Figure~\ref{fig:exp_inference_top_basic_ops} in the response compares the top 5 time-consuming basic operators in inference.
%
For GCN, the \texttt{index} operator used in the prediction layer became the new top 5 time-consuming operator, replacing the \texttt{gather} operator used in the backward phase in training.
%
%    The basic operators related to the edge calculation (\texttt{scatter\_add}, \texttt{index\_select}, and \texttt{mul}) still consumed the majority of the inference time.
%
For GGNN, \texttt{index} operator in the prediction layer also became one of the top 5 time-consuming basic operators.
%
%    Due to the high time complexity of the vertex updating function $\gamma$ in GGNN, the basic operators related to the vertex calculation stage (\texttt{mm} and \texttt{thnn\_fused\_gru\_cell}) still consumed near half of the inference time.
%
For GAT, the \texttt{input\_put\_impl} operator (used in the backward phase) in Figure~\ref{fig:exp_top_basic_ops}(c) was replaced by the \texttt{scatter\_add} operator used in the forward edge calculation in Figure~\ref{fig:exp_inference_top_basic_ops}(c).
%
%    The \texttt{scatter\_add} operator was related to the edge calculation step.
%
For GaAN, the \texttt{mm} operator was replaced by the \texttt{cat} operator used in the vertex updating function.

Although some specific time-consuming operators were different, the performance bottlenecks on the operator level were same for training and inference.
%
The matrix multiplication \texttt{mm} and the element-wise multiplication \texttt{mul} operators were still time-consuming for inference, making GNN inference also {suitable for GPUs}.
%
The basic operators related to the collection and aggregation steps in the edge calculation still consumed non-trivial time in inference.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gaan.pdf}}
    \caption{Top 5 time-consuming basic operators of typical GNNs. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
    \label{fig:exp_inference_top_basic_ops}
\end{figure}
